Using HyperNEAT vs. Parameter Optimization to Create Quadruped Robot Gaits

### 2011-04-13 paper_ecal/introduction.tex
forging

Legged robots have the potential to access many types of terrain
unsuitable for wheeled robots, but doing so requires the creation of a
gait specifying how the robot is to walk.  Gaits may be designed
by one of several methods: either manually by an expert or
via computer learning algorithms.  Learned gaits
offer several advantages over manually designed gaits.  Automatically
learning gaits can save valuable engineering time and can allow
customization of gaits to a particular robot and its unique actuators.
Most importantly, in many cases learned gaits can outperform manually
designed gaits~\citep{hornby2005autonomous, valsalam2008modular}.

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\citep{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\citep{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_whitebg.jpg}) on flat terrain.

@@

A final motivation of this problem is to simply evolve effective gaits
for a physical robot. Because HyperNEAT has performed well in
simulation, it is interesting to test whether it can produce a fast
gait for a physical robot. It is additionally interesting to test how
more traditional gait optimization techniques compete with
evolutionary algorithms when evolving in hardware.

>>>>

Legged robots have the potential to access many types of terrain
unsuitable for wheeled robots, but doing so requires the creation of a
gait specifying how the robot walks.  Such gaits may be designed either manually by an expert or
via computer learning algorithms.  It is advantageous to automatically
learn gaits because doing so can save valuable engineering time and allows
gaits to be customized to the idiosyncrasies of different robots.
Additionally, learned gaits have outperformed engineered gaits in some cases~\citep{hornby2005autonomous, valsalam2008modular}.

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\citep{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\citep{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_whitebg.jpg}).

@@

An additional motivation is to test whether gaits evolved in simulation transfer to reality well, especially those evolved with cutting-edge evolutionary algorithms. Because HyperNEAT gaits performed well in
simulation, it is interesting to test whether HyperNEAT can produce fast
gaits for a physical robot. Such tests help us better understand the real world implications of results reported only in simulation. It is additionally interesting to test how
more traditional gait optimization techniques compete with
evolutionary algorithms when evolving in hardware. A final motivation of this research is simply to evolve effective gaits for a physical robot.

### 2011-04-13 paper_ecal/problemDefinition.tex
forging

For the system evaluated in this paper, we chose to compare only
feedforward gaits generated by both the parameterized methods and
HyperNEAT.  An interesting extension would be to allow gaits to depend
on the measured servo positions, loads, voltage drops, or other
quantities.

>>>>

For the system evaluated in this paper, we chose to compare open-loop gaits generated by both the parameterized methods and
HyperNEAT.  An interesting extension would be to allow closed-loop gaits that depend
on the measured servo positions, loads, voltage drops, or other
quantities.

### 2011-04-13 paper_ecal/relatedWork.tex
forging

Various machine learning techniques have proved to be effective at
generating gaits for legged robots. Kohl and Stone presented a policy
gradient reinforcement learning approach for generating a fast walk on
legged robots \citep{kohl}, which we implemented for comparison. Others
have evolved gaits for legged robots, producing competitive
results \citep{chernova2005evolutionary, hornby2005autonomous, zykov,
  clune2009evolving, clune2011performance, clune2009hybrid,
  clune2009sensitivity, tellez2006evolving, valsalam2008modular}. In
fact, an evolved gait was used in the first commercially-available
version of Sony's AIBO robot~\citep{hornby2005autonomous}. Except for
work with HyperNEAT \citep{clune2009evolving, clune2011performance,
  clune2009hybrid, clune2009sensitivity}, the previous evolutionary
approaches have helped evolution exploit the regularity of the problem
by manually decomposing the task. Experimenters have
to choose which legs should be coordinated, or otherwise facilitate
the coordination of motion. Part of the motivation of this paper
is to see if the space of regularities can be explored automatically,
which has previously performed well \citep{valsalam2008modular}, and to
perform a direct comparison on the same robot to HyperNEAT.

>>>>

Various machine learning techniques have proved to be effective at
generating gaits for legged robots. Kohl and Stone presented a policy
gradient reinforcement learning approach for generating a fast walk on
legged robots \citep{kohl}, which we implemented for comparison. Others
have evolved gaits for legged robots, producing competitive
results \citep{chernova2005evolutionary, hornby2005autonomous, zykov,
  clune2009evolving, clune2011performance, clune2009hybrid,
  clune2009sensitivity, tellez2006evolving, valsalam2008modular}. In
fact, an evolved gait was used in the first commercially-available
version of Sony's AIBO robot~\citep{hornby2005autonomous}. Except for
work with HyperNEAT \citep{clune2009evolving, clune2011performance,
  clune2009hybrid, clune2009sensitivity}, the previous evolutionary
approaches have helped evolution exploit the regularity of the problem
by manually decomposing the task. Experimenters have
to choose which legs should be coordinated, or otherwise facilitate
the coordination of motion. Part of the motivation of this paper
is to see if the space of regularities can be explored more systematically,
which has previously performed well \citep{valsalam2008modular}, and to
compare such an approach to HyperNEAT.
